---
title: "VectorSurv Sample Report using R"
author: "Prepared by Christina and Olivia"
date: "October 25, 2023"
output:
  html_document:
    df_print: paged
geometry: left=1in,right=1in,top=1in,bottom=1in
---
```{r configure-chunks, include=FALSE}
knitr::opts_chunk$set(
  message=FALSE, warning=FALSE, results='asis')

```

```{r working-directory, include=FALSE}
# getwd()
# setwd("/Users/...")

```

```{r load-source-file, include=FALSE}
source("VS_functions.R")
```

```{r retrieve-data-from-API, include=FALSE}
##Uncomment the code below and and run it once. Before moving on, recomment the code. This will prevent R from retrieving your data each time you knit your file. 

# collections = getArthroCollections(2022,2023)
# pools = getPools(2022,2023)
```

```{r write-data-to-csv, include=FALSE}
##Comment out this code after writing it once so it doesn't write each time you knit your file. 

# write.csv2(collections, file="collections_22_23.csv")
#write.csv2(pools, file="pools_22_23.csv")
```

```{r read-data-into-R, echo=FALSE}
##Read your csvs into your R workspace to continue working with them
collections = read.csv("collections_18_23.csv", sep = ",")
pools = read.csv("pools_18_23.csv", sep=",")
```

I can include any text relevant to the report, such as a quick summary of what's happening at this point in the season and I can **bold anything that is important** or maybe *italicize it*. This is using the basics of [Markdown](https://www.markdownguide.org/cheat-sheet/). I can also create specific sections, such as Trapping effort, Abundance, West Nile Activity, etc, and can change the header size.

## Abundance

I can include text anywhere outside of the code chunks and the text will show up in my final file. 


```{r create-abundance-table, echo=FALSE}
abundance2023 <- getAbundance(collections,
             interval = "Week",
             species_list = c("Cx tarsalis", "Cx pipiens"),
             trap_list = "CO2", 
             species_seperate = TRUE)


abundance2023 %>% 
  filter(EPIYEAR=="2023") %>%
  filter(Week>"36") %>%
  kbl(col.names = c("Disease Year","Week","Species" ,"Count","Trap Events", "Mosquitoes per trap night")) %>%
  kable_styling(bootstrap_options = "striped",
                              font_size = 14, 
                html_font = "Cambria", full_width = F)%>%  
  footnote(general = "Table X: Weekly Abundance Calculation for Cx. tarsalis, pipiens in CO2 traps", general_title = "")

```

```{r 2023-abundance-plot, echo=FALSE, fig.width=10}
abundance2023 %>%  
  filter(EPIYEAR=="2023") %>%
  ggplot(aes(x=Week,
             y= Abundance)) +
  geom_point()+ 
  geom_line() + 
  facet_wrap(~species_display_name)+
  labs(title="2023 Abundance",y = "Mosquitoes per trap night")+
  theme(
    plot.title = element_text(size = 18),
     axis.title = element_text(size = 14)
  )

```



