---
title: "Tutortial Documentation"
output: html_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
```


# Introduction

### Source files, directory
```{r}

#To get current directory
getwd()

#'VS_functions.R' must be in the same directory as the .rmd files which access its functions, you can set the working directory 

#Set below to the location of the downloaded files on your machine
setwd("/Users/Christina/Desktop/R-tutorial")

#loads relevant packages and functions
source("VS_functions.R")

```

# Retriving Data

**getArthroCollections(...)**
*Description*

getArthroCollections(...) obtains data on a year range (start_year, end_year) and agency_id. It prompts the user for their gateway username and password before retrieving the associated data.

*Usage*

getArthroCollections(start_year, end_year, agency_code)

*Arguments*

- start_year: Beginning of year range 
- end_year: End of year range
- agency_code: Agency ID number 


```{r}
#Example
collections = getArthroCollections(2022,2023, 55)
```


```{r}


#Example
#pools = getPools(2022,2023, 55)
```

## Write Data to file

 write.csv(...)

 read.csv(...)

```{r}

```


## Basic subseting and filtering

```{r}
#Subset using column names or index
colnames(collections)

collections[c("collection_date","species_display_name","num_count")]

#We can filter using column names


filter(collections,trap_nights==1)

```

# Calculations


## Abundance
getAbundance()

```{r}
```

## Abundance Anomaly (comparison to 5 year average)

getAbundanceAnomaly()

```{r}
```

## Infection Rate

getInfectionRate()

```{r}
```

## Vector Index

getVectorIndex()

```{r}
```

# Tables

getTrapTypeTally()

```{r}
```


getPoolsComparisionTable()

```{r}
```



# Charts

#Requires
ProcessAbunAnom()
plotInfectionRate()
```{r}


```

